---
title: "Practicum 2"
output: html_notebook
---

## Support Functions

#### Function: parseJournal
Returns the data frame for journal when an xml object is passed to this function.

```{r parseJournal}
parseJournal <- function(journalItem) {
  children <- xmlChildren(journalItem)
  
  title <- xmlValue(children$Title)
  issn <- xmlValue(children$ISSN)
  
  return(data.frame(
    title,
    issn
  ))
}
```

#### Function: parseAuthor
Returns the data frame for the author from an xml object which is passed to this function.
```{r}
parseAuthor <- function(authorItem) {
  children <- xmlChildren(authorItem)
  lname <- xmlValue(children$LastName)
  fname <- xmlValue(children$ForeName)
  initial <- xmlValue(children$Initials)
  affiliation <- xmlValue(children$Affiliation)
  
  return(data.frame(
    lname,
    fname,
    initial,
    affiliation
  ))
}
```

#### Function: rowExists
Checks if the row exists in th given dataframe and returns the key of the existing row else returns 0.

```{r rowExists}
rowExists <- function(df, row) {
  r <- nrow(df)
  
  if(r == 0){
    return(0)
  }
  for(a in 1:r){
    if(all(df[a,]==row[1,])) {
      return(a)
    }
  }
  return(0)
}
```

```{r}
cleanString <- function(dirtyString){
  cleanedString <- str_replace_all(dirtyString, "[^[:alnum:]]", "")
  return(cleanedString) 
}
```


```{r}
# Importing required libraries
if("RMySQL" %in% rownames(installed.packages()) == FALSE) {
  install.packages("RMySQL")
}
library(RMySQL)

if("XML" %in% rownames(installed.packages()) == FALSE) {
  install.packages("XML")
}
library(XML)
library(stringr)
```

```{r}
# Connecting to Database

#Connection setting
db_user <- 'sudo'
db_password <- 'dbmsp2neu'
db_name <- 'med_db'
db_host <- 'db-practicum-2.cfm4y0dsbhpu.us-east-2.rds.amazonaws.com'
db_port <- 3306

#Connect to the DB
dbcon <- dbConnect(MySQL(), user = db_user, password = db_password, dbname = db_name, host = db_host, port = db_port)
```

```{r}
# Importing the XML file
path <- ""
xmlFile <- "pubmed_sample.xml"
fp <- paste0(path,xmlFile)
xmlDOM <- xmlParse(fp)
```

```{r}
# get the root node of the DOM tree
root <- xmlRoot(xmlDOM)
# get number of children of root 
sizeRoot <- xmlSize(root)
```

```{r}
# create various data frames to hold data;
author <- data.frame (id = integer(),
                      lname = character(),
                      fname = character(),
                      initial = character(),
                      affiliation = character(),
                      stringsAsFactors = F)

journal <- data.frame (id = integer(),
                       title = character(),
                       issn = character(),
                       stringsAsFactors = F)

journalIssue <- data.frame (id = integer(),
                            volume = integer(),
                            issue = integer(),
                            journal_id = integer(),
                            pub_date = character(),
                            stringsAsFactors = F)

article <- data.frame (id = integer(),
                       title = character(),
                       pub_date = character(),
                       journal_id = integer(),
                       pub_model = character(),
                       stringsAsFactors = F)


articleAuthorMapping <- data.frame (id = integer(),
                                    author_id = integer(),
                                    article_id = integer(),
                                    stringsAsFactors = F)


history <- data.frame (id = integer(),
                       date_time = character(),
                       type = character(),
                       article_id = integer(),
                       stringsAsFactors = F)

```

## Parsing XML data to dataframes

#### Parsing &lt;Journal&gt; to dataframe
```{r}
journalListXpath <- "/PubmedArticleSet/PubmedArticle/MedlineCitation/Article/Journal"
journalList <- xpathSApply(xmlDOM, journalListXpath)

counter <- 0
for(i in 1:sizeRoot) {
  item <- parseJournal(journalList[[i]])
  
  if(rowExists(journal[, 2:ncol(journal)], item) == 0){
    journal[counter,"id"] = counter
    journal[counter,"title"] = item$title
    journal[counter,"issn"] = item$issn 
    
    counter <- counter + 1
  }
}
```

#### Parsing &lt;JournalIssue&gt; to dataframe
```{r}
journalIssueXPath <- "/PubmedArticleSet/PubmedArticle/MedlineCitation/Article/Journal/JournalIssue"
journalIssueList <- xpathSApply(xmlDOM, journalIssueXPath)

for(i in 1:sizeRoot) {
  
}
```

#### Parsing all Authors to dataframe
```{r}
authorListXpath <- "/PubmedArticleSet/PubmedArticle/MedlineCitation/Article/AuthorList/Author"

authorList <- xpathSApply(xmlDOM, authorListXpath)
counter <- 1
for(i in 1:xmlSize(authorList)){
  item <- parseAuthor(authorList[[i]])
  item$lname = cleanString(item$lname)
  item$fname = cleanString(item$fname)
  item$initial = cleanString(item$initial)
  
  if(rowExists(author[, 2:(ncol(author)-1)], item[,1:(ncol(item)-1)]) == 0){
    author[counter,"id"] = counter
    author[counter,"lname"] = item$lname
    author[counter,"fname"] = item$fname
    author[counter,"initial"] = item$initial
    author[counter,"affiliation"] = item$affiliation
    counter <- counter + 1
  }
}
```

```{r}
# Disconnect the database
dbDisconnect(dbcon)
```


